On centrality of orthogonal $\Kt$

### 2016-04-21 paper.tex
Writing introduction

\newcommand{\Max}{\mathop{\mathrm{Max}}\nolimits}
\newcommand{\proj}{\mathop{\mathrm{proj}}\nolimits}
\newcommand{\Card}{\mathop{\mathrm{Card}}\nolimits}
\newcommand{\Ker}{\mathop{\mathrm{Ker}}\nolimits}
\newcommand{\Cent}{\mathop{\mathrm{Cent}}\nolimits}
\newcommand{\E}{{\mathrm{E}}}
\newcommand{\Um}{\mathop{\mathrm{Um}}\nolimits}
\newcommand{\St}{\mathop{\mathrm{St}}\nolimits}
\newcommand{\Sp}{\mathop{\mathrm{Sp}}\nolimits}
\newcommand{\Ep}{\mathop{\mathrm{Ep}}\nolimits}
\newcommand{\GL}{\mathop{\mathrm{GL}}\nolimits}
\newcommand{\SL}{\mathop{\mathrm{SL}}\nolimits}
\newcommand{\Kt}{\mathop{\mathrm{K_2}}\nolimits}
\newcommand{\Ko}{\mathop{\mathrm{K_1}}\nolimits}
\newcommand{\Ho}{\mathop{\mathrm{H_1}}\nolimits}
\newcommand{\Ht}{\mathop{\mathrm{H_2}}\nolimits}
\newcommand{\epi}{\twoheadrightarrow}
\newcommand{\sgn}{\mathrm{sgn}}
\newcommand{\eps}[1]{\varepsilon_{#1}}
\newcommand{\sign}[1]{\mathrm{sign}(#1)}
\newcommand{\lan}{\langle}
\newcommand{\ran}{\rangle}
\newcommand{\inv}{^{-1}}
\newcommand{\ur}[1]{\!\,^{(#1)}U_1}
\newcommand{\ps}[1]{\!\,^{(#1)}\!P_1}
\newcommand{\ls}[1]{\!\,^{(#1)}\!L_1}

@@

\section*{Inroduction}

@@

The local--global principle for the elementary group proven by Suslin~\cite{Sus} is a statement that an elementary matrix over a polynomial ring $g\in\E(n,\,R[X])$ becomes trivial after the evaluation ``$X=0$''
$$
g\in\Ker\big(\E(n,\,R[X])\rightarrow\E(n,\,R)\big)
$$
and after localisations
$$
g_{\mathfrak m}=1\in\E(n,\,R_{\mathfrak m}[X])
$$
for all maximal ideals $\mathfrak m$ of a base ring $R$, if and only if it is trivial from the very begining
$$
g=1\in\E(n,\,R[X]).
$$
An analogous statement about Steinberg groups was proven by Tulenbaev for $n\geq5$~\cite[Th.~2.1]{Tul}. I.e., for $n\geq5$ and $g\in\St(n,\,R[X])$ such that 
$$g(0)=1\in\St(n,\,R)\ \text{ and }\ g_{\mathfrak m}=1\in\St(n,\,R_{\mathfrak m}[X])$$
 for all $\mathfrak m\in\Max R$ holds $g=1$.

@@

>>>>

\newcommand{\Max}{\mathop{\mathrm{Max}}\nolimits}
\newcommand{\proj}{\mathop{\mathrm{proj}}\nolimits}
\newcommand{\Card}{\mathop{\mathrm{Card}}\nolimits}
\newcommand{\Ker}{\mathop{\mathrm{Ker}}\nolimits}
\newcommand{\Cent}{\mathop{\mathrm{Cent}}\nolimits}
\newcommand{\E}{{\mathrm{E}}}
\newcommand{\Um}{\mathop{\mathrm{Um}}\nolimits}
\newcommand{\St}{\mathop{\mathrm{St}}\nolimits}
\newcommand{\Sp}{\mathop{\mathrm{Sp}}\nolimits}
\newcommand{\Ep}{\mathop{\mathrm{Ep}}\nolimits}
\newcommand{\GL}{\mathop{\mathrm{GL}}\nolimits}
\newcommand{\SL}{\mathop{\mathrm{SL}}\nolimits}
\newcommand{\Kt}{\mathop{\mathrm{K_2}}\nolimits}
\newcommand{\Ko}{\mathop{\mathrm{K_1}}\nolimits}
\newcommand{\Ho}{\mathop{\mathrm{H_1}}\nolimits}
\newcommand{\Ht}{\mathop{\mathrm{H_2}}\nolimits}
\newcommand{\epi}{\twoheadrightarrow}
\newcommand{\sgn}{\mathrm{sgn}}
\newcommand{\eps}[1]{\varepsilon_{#1}}
\newcommand{\sign}[1]{\mathrm{sign}(#1)}
\newcommand{\lan}{\langle}
\newcommand{\ran}{\rangle}
\newcommand{\inv}{^{-1}}
\newcommand{\ur}[1]{\!\,^{(#1)}U_1}
\newcommand{\ps}[1]{\!\,^{(#1)}\!P_1}
\newcommand{\ls}[1]{\!\,^{(#1)}\!L_1}
\newcommand{\rA}{\mathsf{A}}
\newcommand{\rB}{\mathsf{B}}
\newcommand{\rC}{\mathsf{C}} 
\newcommand{\rD}{\mathsf{D}} 
\newcommand{\rE}{\mathsf{E}}
\newcommand{\rF}{\mathsf{F}}
\newcommand{\rG}{\mathsf{G}}

@@

\section*{Inroduction}
The main goal of the present paper is to obtain a uniform proof of the local-global principle for $\Kt$ and centrality of $\Kt$ for Chevalley groups of type $ADE$.

Recall that for each reduced root system $\Phi$ and a commutative ring $R$ one can construct a split simple group $G(\Phi, R)$ called \emph{Chevalley group}.
For $\Phi$ of rank $\geq 2$ the abstract group $G(\Phi, R)$ contains a large normal subgroup $\E(\Phi, R)$ generated by the elementary root unipotents $t_\alpha(\xi)$, $\xi\in R$, $\alpha\in \Phi$.
The quotient group $\Ko(\Phi, R)=G(\Phi, R)/\E(\Phi, R)$ is called the \emph{unstable $\Ko$-functor}.
This functor shares many properties with its stable counterpart $\mathrm{SK}_1(R) = \lim\limits_{l\to\infty}\Ko(\rA_\ell, R)$ which, in turn, is an essential direct summand of $\Ko(R)$.

Similarly to the definition of the algebraic $\Kt$-functor, one can define $\Kt(\Phi, R)$ as the kernel of the canonical projection $\varphi\colon\St(\Phi, R)\twoheadrightarrow \E(\Phi, R)$.
Here $\St(\Phi, R)$ stands for the \emph{Steinberg group} of type $\Phi$, i.\,e. the group defined by formal generators $x_\alpha(\xi)$ modeling root unipotents $t_\alpha(\xi)$
and the explicit list of relations called Steinberg relations. The projection $\varphi$ is defined by the identity $\varphi(x_\alpha(\xi))=t_\alpha(\xi)$.

It is classically known that $\Kt(\Phi, R)$ is a central subgroup of $\St(\Phi, R)$ if $R$ is commutative local ring and $\Phi$ has rank $\geq 2$ (see \cite{Stein}).
However, to prove that centrality of $\Kt$ holds for \emph{arbitrary} $R$ is much harder.
The first result in this direction was obtained in 1977 by W. van~der~Kallen for $\Phi=\rA_\ell$, $\ell\geq 3$.
Using a similar method the first-named author established centrality of $\Kt(\rC_\ell, R)$, see \cite{Lav}.
On the other hand, the second-named author has obtained centrality of $\Kt(\rE_\ell, R)$ using a different method, see \cite{SCh}.
In turn, our main result is the following theorem.
\begin{tm}\label{tm:main} Let $R$ be an arbitrary commutative ring.
 If $\Phi$ is an irreducible root system of type $\Phi=\rA_\ell, \rD_\ell, \rE_\ell$, $\ell\geq 3$ then the following statements are true:
 \begin{enumerate}
  \item The group $\St(\Phi, R)$ satisfies local-global principle for $\Kt$, i.\,e. an element $g\in\St(n,\,R[X])$ such that $g(0)=1\in\St(n,\,R)$
  is trivial if and only if $g_M=1\in\St(n,\,R_M[X])$ for all maximal ideals $M\triangleleft R$.
  \item The group $\Kt(\Phi, R)$ is contained in the centre of $\St(\Phi, R)$.  
 \end{enumerate}
\end{tm}
The special case $\Phi=\rA_\ell$, $\ell\geq 3$ of the first statement of the above theorem was proven for the first time by Tulenbaev~\cite[Th.~2.1]{Tul}, while the case $\Phi=\rE_\ell$ was studied in~\cite{SCh}.
Thus, the new cases in \cref{tm:main} are $\Phi=\rA_3$ and $\Phi=\rD_\ell$, $\ell\geq 4$.

@@

The local--global principle for the elementary group proven by Suslin~\cite{Sus} is a statement that an elementary matrix over a polynomial ring $g\in\E(n,\,R[X])$ becomes trivial after the evaluation ``$X=0$''
$$
g\in\Ker\big(\E(n,\,R[X])\rightarrow\E(n,\,R)\big)
$$
and after localisations
$$
g_{\mathfrak m}=1\in\E(n,\,R_{\mathfrak m}[X])
$$
for all maximal ideals $\mathfrak m$ of a base ring $R$, if and only if it is trivial from the very begining
$$
g=1\in\E(n,\,R[X]).
$$

@@

\section{Yet another presentation for $\St(4, R)$.}
